---
title: "OPSD GERMANY DATA SET OF ELECTRICITY CONSUMPTION"
author: "anit"
date: "2022-08-29"
output:
  pdf_document: default
  html_document: default
---

## The DATA SET : Open Power Systems Data

We'll be working with daily time series of Open Power System Data (OPSD)
for Germany, which has been rapidly expanding its renewable energy
production in recent years. The data set includes country-wide totals of
electricity consumption, wind power production, and solar power
production for 2006-2017. link for the data [click
here](https://raw.githubusercontent.com/jenfly/opsd/master/opsd_germany_daily.csv)

### The columns of the data file are:

-   Date --- The date (yyyy-mm-dd format)
-   Consumption --- Electricity consumption in GWh
-   Wind --- Wind power production in GWh
-   Solar --- Solar power production in GWh
-   Wind+Solar --- Sum of wind and solar power production in GWh

##We will explore how electricity consumption and production in Germany
have varied over time :

-   When is electricity consumption typically highest and lowest?

-   How do wind and solar power production vary with seasons of the
    year?

-   What are the long-term trends in electricity consumption, solar
    power, and wind power?

## CREATING THE DATAFRAME FROM DATA

```{r dataframe, message=FALSE, warning=FALSE, paged.print=FALSE}
library(readxl)
library(tidyverse)
mydata <- read_excel("E:/R/OPSV_GERMANY.xlsx")
View(mydata)
```

#looking at head and tail of dataframe

```{r}
head(mydata)
tail(mydata)
```

#adding year date month from dataset

```{r}
x<-mydata$Date

year<-as.numeric(format(x,"%Y"))
View(year)

month <- as.numeric(format(x,"%m"))

day<-as.numeric(format(x,"%d"))

head(month)
```

#add these 3 colom in dataset

```{r}
newdata<- cbind(mydata,year,month,day)
head(newdata)
```

#changing data types for usage in plots

```{r}
newdata$Date<- as.Date(newdata$Date)
newdata$Consumption<-as.numeric(newdata$Consumption)
newdata$Wind<-as.numeric(newdata$Wind)
newdata$Solar<-as.numeric(newdata$Solar)
newdata$`Wind+Solar`<-as.numeric(newdata$`Wind+Solar`)
summary(newdata)
```

#comsumption over years

```{r}
library(ggplot2)
ggplot(data=newdata)+geom_line(mapping=aes(x=Date,y=Consumption),colour="blue") + labs(title = "Consumption and date relation")+ labs(subtitle = "diff year different color") +labs(y="Consumption in GWh")+labs(x="date")

ggplot(data=newdata)+geom_line(mapping=aes(x=Date,y=Wind),colour="red") + labs(title = "Wind power production over years")+ labs(subtitle = "diff year different color") +labs(y="Wind energy in GWh")+labs(x="date")
ggplot(data=newdata)+geom_line(mapping=aes(x=Date,y=Solar),colour = "green") + labs(title = "Solar production over years")+ labs(subtitle = "diff year different color") +labs(y="Energy in GWh")+labs(x="date")
```

We don't get not much of data from this plot as all are line points We
then use the point plot.

```{r}
library(ggpubr)
ggplot(data=newdata)+geom_point(mapping=aes(x=Date,y=Consumption,color=year),colour="orange") + labs(title = "Consumption and date relation")+labs(y="Consumption in GWh")+labs(x="date")
ggplot(data=newdata)+geom_point(mapping=aes(x=Date,y=Wind),colour="red") + labs(title = "Wind power production over years")  +labs(y="Wind energy in GWh")+labs(x="date")

ggplot(data=newdata)+geom_point(mapping=aes(x=Date,y=Solar),color="green") + labs(title = "Solar production over years")  +labs(y="Energy in GWh")+labs(x="date")


```

### To see trend in a year let's see a part of 2 years from the dataset (using filter)

```{r}
newdata %>% filter(Date>="2011-12-04"& Date<="2013-02-04")  %>%  ggplot()+geom_line(mapping=aes(x=Date,y=Consumption),colour="blue") +labs(title = "electricity consumption between 2011-12-04 and 2013-02-04 ")

newdata %>% filter(Date>="2006-12-04"& Date<="2008-02-04")  %>%  ggplot()+geom_line(mapping=aes(x=Date,y=Consumption),colour="blue") +labs(title = "electricity consumption between 2006-12-04 and 2008-02-04 ")
```

### From the consumption data between 2011-12-04 and 2013-02-04 and between 2006-12-04 and 2008-02-04 It can be noted:

-   In summer the comsumption in both timeline decreases

-   during Winter the usage of electricity increases.(mostly due to
    heater usage)

-   there are significant drop of uasge during winter holiday (ie during
    Christmas and new year eve)

-   from the below point plot of consumption history between 2011-12-04
    and 2016-02-04 it is seen the consumption splits into 2 clusters.
    one around 1400 GWh and another around 1100 GWh and oscillates
    around these two values.(during weekands and weekdays )

```{r}
newdata %>% filter(Date>="2011-12-04"& Date<="2016-02-04")  %>%  ggplot()+geom_point(mapping=aes(x=Date,y=Consumption),colour="blue") +labs(title = "electricity consumption between 2011-12-04 and 2016-02-04 ")+ labs(subtitle = "to find relevant pattern in consumption")

```

#about the solar power production variation lets take a different
timespan

```{r}
newdata %>% filter(Date>="2011-12-04"& Date<="2014-02-04")  %>%  ggplot()+geom_point(mapping=aes(x=Date,y=Solar),colour="blue") +labs(title = "electricity production by solar between 2011-12-04 and 2014-02-04 ")

```

###conclusion: Solar power production is highest in time of Summer of
the season(june-july) and lowest in the winter time of the year.

### to find similar pattern in Wind energy production

```{r}
newdata %>% filter(Date>="2011-12-04"& Date<="2013-02-04")  %>%  ggplot()+geom_line(mapping=aes(x=Date,y=Wind),colour="blue") +labs(title = "electricity production by Wind between 2011-12-04 and 2013-02-04 ")+labs(subtitle = "line plot")
newdata %>% filter(Date>="2011-12-04"& Date<="2013-02-04")  %>%  ggplot()+geom_point(mapping=aes(x=Date,y=Wind),colour="blue") +labs(title = "electricity production by Wind between 2011-12-04 and 2013-02-04 ") +  labs(subtitle = "point plot")

```

#### conclusion:

-   Wind power production maxes out in Winter min in Summer

moreover from the whole data plot of Wind production:

```{r}
ggplot(data=newdata)+geom_line(mapping=aes(x=Date,y=Wind),colour="red") + labs(title = "Wind power production over years")  +labs(y="Wind energy in GWh")+labs(x="date")
```

**Over the years starting from 2010 the amount of energy production
gradually increases**

##CONCLUSIONS: \* there are periodicity happening in all three
parameters CONSUMPTION,WIND POWER PRODUCTION,SOLAR POWER PRODUCTION in
span of a year. \*all three time series clearly exhibit
periodicity---often referred to as seasonality in time series
analysis---in which a pattern repeats again and again at regular time
intervals. The Consumption, Solar, and Wind time series oscillate
between high and low values on a yearly time scale, corresponding with
the seasonal changes in weather over the year.

#### let's further look into monhly data and see how data varies over days of the month

```{r}
newdata %>% filter(Date>="2017-01-04"& Date<="2018-01-04")  %>%  ggplot()+geom_line(mapping=aes(x=Date,y=Consumption),colour="blue") +labs(title = "electricity consumption between 2017-01-04 and 2018-01-04 ") +labs(subtitle = "a year's data of consumption in GWh")


```

```{r}
newdata %>% filter(Date>="2017-01-01"& Date<="2017-03-04")  %>%  ggplot()+geom_line(mapping=aes(x=Date,y=Consumption),colour="blue") +geom_point(mapping=aes(x=Date,y=Consumption),colour="red") +labs(title = "electricity consumption between 2017-01-04 and 2017-03-04 ") +labs(subtitle = "2 month's  data of consumption in GWh")
```

**As we suspected, consumption is highest on weekdays and lowest on
weekends.**

##Seasonality

Next, let's further explore the seasonality of our data with box plots

```{r}
h1<-ggplot(data=newdata) +geom_boxplot(aes(group=month,x=month, y= Consumption),fill="yellow")+labs(title = "Consumption") +scale_x_continuous(breaks=c(1:12))
  
h2<-ggplot(data=newdata) +geom_boxplot(aes(group=month,x=month, y= Wind),fill="yellow")+scale_x_continuous(breaks=c(1:12)) 
h3<-ggplot(data=newdata) +geom_boxplot(aes(group=month,x=month, y= Solar),fill="orange")+ labs(x="month")+scale_x_continuous(breaks=c(1:12))
ggarrange(h1,h2,h3,ncol=1,nrow=3)
```

These box plots confirm the yearly seasonality that we saw in earlier
plots and provide some additional insights: \* electricity consumption
is generally higher in winter and lower in summer, the median and lower
two quartiles are lower in December and January compared to other
months. likely due to businesses being closed over the holidays. We saw
this in the time series for the year 2017, that for some points in JAN
and DEC there are consumptions much lower with respect to other points.

-   solar and wind power production both exhibit a yearly seasonality,
    the wind power distributions have many more outliers, reflecting the
    effects of occasional extreme wind speeds associated with storms and
    other transient weather conditions. ***points exiting the range of
    Q3+1.5 of IQR***

###Next, let's group the electricity consumption time series by day of
the week, to explore weekly seasonality.

For that we need to enter the weekdays of the given data set

```{r}
weekday<-weekdays(newdata$Date)
newdata<-cbind(newdata,weekday)

newdata$weekday <- factor(newdata$weekday, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
```

```{r}
newdata %>% filter(year=="2017") %>% ggplot() +geom_boxplot(aes(group=weekday,x=weekday, y= Consumption,fill=weekday))+labs(title = "box plot of consumption data by Weekdays of the Year 2017")
```

As expected, electricity consumption is significantly higher on weekdays
than on weekends. The low outliers on weekdays are presumably during
holidays.

```{r}
library(dplyr)
library(zoo)
library(dplyr)
library(zoo)
test1data<- newdata  %>% arrange(day)%>% arrange(month) %>% arrange(year)  %>%  mutate(avg_of_7 =rollmean(Consumption,k=7,fill = NA) ) 
View(test1data)
```

```{r}
test2data<- test1data  %>% arrange(day)%>% arrange(month) %>% arrange(year)  %>%  mutate(avg_of_3 =rollmean(Consumption,k=3,fill = NA)) 
View(test2data)
```

```{r}
testdata<- test2data  %>% arrange(day)%>% arrange(month) %>% arrange(year)  %>%  mutate(avg_of_365 =rollmean(Consumption,k=365,fill = NA)) 
View(testdata)
```

### trend of rollMeans of 7,3,365 days

```{r}
data_new<- data.frame(x=testdata$Date,y=c(testdata$avg_of_7,testdata$avg_of_3,testdata$avg_of_365),group=c(rep("avg_of_7",nrow(testdata)),rep("avg_of_3",nrow(testdata)),rep("avg_of_365",nrow(testdata))))

ggplot(data=data_new) +geom_line(mapping = aes(x,y,col=group))+scale_colour_manual(values=c("avg_of_7"="green", "avg_of_3"="blue", "avg_of_365"="white"), 
                                                                                   labels=c("avg_of_7", "avg_of_3", "avg_of_365"))+xlab("Date")+ylab("GWh")


```

Conclusion from the trend: \* the 3 and 7 day mean data of consumption
is still fluctuating in nature \* the 365 days rollMean/continous mean
is still having almost similar range of results throughout the data
cycle
